### 2022-03-31 R/DeaminationAnalysis.rmd
Create DeaminationAnalysis.rmd
An R notebook for displaying results for the deamination analysis.

---
title: "R Notebook"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)
opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5)
if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}
```

```{r acquire_and_format_data}
deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R', "DeaminationAnalysis.R"))

# Read in the relevant data
CPDMismatchDataFilePaths = list(hr1 = file.path(deaminationDataDirectory, "NHF1_CPD_1h_Rep1_mismatches.bed"),
                                hr24 = file.path(deaminationDataDirectory, "NHF1_CPD_24h_Rep1_mismatches.bed"),
                                hr48 = file.path(deaminationDataDirectory, "NHF1_CPD_48h_Rep1_mismatches.bed"))

sixFourMismatchDataFilePaths = list(hr1 = file.path(deaminationDataDirectory, "NHF1_6-4-PP_1h_Rep1_mismatches.bed"))

CPDMismatchData = lapply(CPDMismatchDataFilePaths, fread)
sixFourMismatchData = lapply(sixFourMismatchDataFilePaths, fread)

# Simplify the data
CPDMismatchData = lapply(CPDMismatchData, function(x) simplifyTable(filterResults(x)))
sixFourMismatchData = lapply(sixFourMismatchData, function(x) simplifyTable(filterResults(x)))

```

```{r plot_mismatch_type_frequency}

plotMismatchTypeFrequencies(CPDMismatchData$hr1, title = "CPD 1hr Mismatch Type Frequency")
plotMismatchTypeFrequencies(CPDMismatchData$hr24, title = "CPD 24hr Mismatch Type Frequency")
plotMismatchTypeFrequencies(CPDMismatchData$hr48, title = "CPD 48hr Mismatch Type Frequency")

plotMismatchTypeFrequencies(sixFourMismatchData$hr1, title = "6-4 1hr Mismatch Type Frequency")

```

```{r plot_mismatch_pos_frequency}

plotMismatchPositionFrequencies(CPDMismatchData$hr1, includedTypes = c("C>T"), 
                                title = "CPD 1hr C>T Mismatch Positions")
plotMismatchPositionFrequencies(CPDMismatchData$hr1, omittedTypes = c("C>T"), 
                                title = "CPD 1hr !C>T Mismatch Positions")

plotMismatchPositionFrequencies(CPDMismatchData$hr24, includedTypes = c("C>T"), 
                                title = "CPD 24hr C>T Mismatch Positions")
plotMismatchPositionFrequencies(CPDMismatchData$hr1, omittedTypes = c("C>T"), 
                                title = "CPD 24hr !C>T Mismatch Positions")

plotMismatchPositionFrequencies(CPDMismatchData$hr48, includedTypes = c("C>T"), 
                                title = "CPD 48hr C>T Mismatch Positions")
plotMismatchPositionFrequencies(CPDMismatchData$hr48, omittedTypes = c("C>T"), 
                                title = "CPD 48hr !C>T Mismatch Positions")

plotMismatchPositionFrequencies(sixFourMismatchData$hr1, includedTypes = c("C>T"), 
                                title = "6-4 1hr C>T Mismatch Positions")
plotMismatchPositionFrequencies(sixFourMismatchData$hr1, omittedTypes = c("C>T"), 
                                title = "6-4 1hr !C>T Mismatch Positions")

```