### 2022-05-03 R/HumanDeaminationReadLengthAnalysis.rmd

---
title: "Human Deamination Read Length Analysis"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)
opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5, 
               fig.path = paste0(file.path("deamination_analysis_output","human_analysis_figures"),
                                 .Platform$file.sep))
if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", "human")
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R', "DeaminationAnalysis.R"))
```

```{r get_file_paths}
# Get file paths for the mismatch by read data
CPDMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
CPDMismatchDataFilePaths$rep1 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_1h_Rep1_mismatches_by_read.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_8h_Rep1_mismatches_by_read.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_24h_Rep1_mismatches_by_read.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_48h_Rep1_mismatches_by_read.bed")
)
CPDMismatchDataFilePaths$rep2 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_1h_Rep2_mismatches_by_read.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_8h_Rep2_mismatches_by_read.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_24h_Rep2_mismatches_by_read.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_48h_Rep2_mismatches_by_read.bed")
)

sixFourMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
sixFourMismatchDataFilePaths$rep1 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_5min_Rep1_mismatches_by_read.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_20min_Rep1_mismatches_by_read.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_1h_Rep1_mismatches_by_read.bed")
)
sixFourMismatchDataFilePaths$rep2 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_5min_Rep2_mismatches_by_read.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_20min_Rep2_mismatches_by_read.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_1h_Rep2_mismatches_by_read.bed")
)
```

```{r acquire_and_format_data}
CPDMismatchData = lapply(CPDMismatchDataFilePaths, function(x) lapply(x, fread))
sixFourMismatchData = lapply(sixFourMismatchDataFilePaths, function(x) lapply(x, fread))

# Create a subset of single mismatch C>T reads.
CPDSingleCToTMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))
sixFourSingleCToTMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))

# Simplify the data
CPDSingleCToTMismatchData = lapply(CPDSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE)
))
sixFourSingleCToTMismatchData = lapply(sixFourSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE)
))

```

```{r filter_by_mismatch_pos_and_read_length}
CPDPeakCToTMismatchData = lapply(CPDSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS)
})
CPDFirstQPeakCToTMismatchData = lapply(CPDPeakCToTMismatchData, function(x) {
  lapply(x, getFirstPositionQuartile)
})
CPDLastQPeakCToTMismatchData = lapply(CPDPeakCToTMismatchData, function(x) {
  lapply(x, getLastPositionQuartile)
})

sixFourPeakCToTMismatchData = lapply(sixFourSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS)
})
sixFourFirstQPeakCToTMismatchData = lapply(sixFourPeakCToTMismatchData, function(x) {
  lapply(x, getFirstPositionQuartile)
})
sixFourLastQPeakCToTMismatchData = lapply(sixFourPeakCToTMismatchData, function(x) {
  lapply(x, getLastPositionQuartile)
})
```

```{r generate_nucleotide_frequency_data}
CPDSingleCToTNucFreq = lapply(CPDSingleCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
CPDFirstQPeakCToTNucFreq = lapply(CPDFirstQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
CPDLastQPeakCToTNucFreq = lapply(CPDLastQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)

sixFourSingleCToTNucFreq = lapply(sixFourSingleCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
sixFourFirstQPeakCToTNucFreq = lapply(sixFourFirstQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
sixFourLastQPeakCToTNucFreq = lapply(sixFourLastQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
```

```{r purine_vs_pyrimidine_frequencies_across_read_length, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDSingleCToTNucFreq$rep1, title = "CPD Rep1 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDSingleCToTNucFreq$rep2, title = "CPD Rep2 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourSingleCToTNucFreq$rep1, title = "6-4 Rep1 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourSingleCToTNucFreq$rep2, title = "6-4 Rep2 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```

```{r purine_vs_pyrimidine_frequencies_first_quartile_mismatch_pos, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDFirstQPeakCToTNucFreq$rep1, title = "CPD Rep1 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDFirstQPeakCToTNucFreq$rep2, title = "CPD Rep2 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourFirstQPeakCToTNucFreq$rep1, title = "6-4 Rep1 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourFirstQPeakCToTNucFreq$rep2, title = "6-4 Rep2 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```

```{r purine_vs_pyrimidine_frequencies_last_quartile_mismatch_pos, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDLastQPeakCToTNucFreq$rep1, title = "CPD Rep1 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDLastQPeakCToTNucFreq$rep2, title = "CPD Rep2 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourLastQPeakCToTNucFreq$rep1, title = "6-4 Rep1 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourLastQPeakCToTNucFreq$rep2, title = "6-4 Rep2 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```