### 2022-05-06 R/HumanDeaminationReadLengthAnalysis.rmd
Update HumanDeaminationReadLengthAnalysis.rmd

---
title: "Human Deamination Read Length Analysis"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)
opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5, 
               fig.path = paste0(file.path("deamination_analysis_output","human_analysis_figures"),
                                 .Platform$file.sep))
if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", "human")
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R', "DeaminationAnalysis.R"))
```

```{r get_file_paths_1}
# Get file paths for the mismatch by read data
CPDMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
CPDMismatchDataFilePaths$rep1 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_1h_Rep1_mismatches_by_read.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_8h_Rep1_mismatches_by_read.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_24h_Rep1_mismatches_by_read.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_48h_Rep1_mismatches_by_read.bed")
)
CPDMismatchDataFilePaths$rep2 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_1h_Rep2_mismatches_by_read.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_8h_Rep2_mismatches_by_read.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_24h_Rep2_mismatches_by_read.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_CPD_48h_Rep2_mismatches_by_read.bed")
)

sixFourMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
sixFourMismatchDataFilePaths$rep1 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_5min_Rep1_mismatches_by_read.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_20min_Rep1_mismatches_by_read.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_1h_Rep1_mismatches_by_read.bed")
)
sixFourMismatchDataFilePaths$rep2 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_5min_Rep2_mismatches_by_read.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_20min_Rep2_mismatches_by_read.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read", "NHF1_6-4-PP_1h_Rep2_mismatches_by_read.bed")
)
```

```{r acquire_and_format_data_1}
CPDMismatchData = lapply(CPDMismatchDataFilePaths, function(x) lapply(x, fread))
sixFourMismatchData = lapply(sixFourMismatchDataFilePaths, function(x) lapply(x, fread))

# Create a subset of single mismatch C>T reads.
CPDSingleCToTMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))
sixFourSingleCToTMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))

# Simplify the data
CPDSingleCToTMismatchData = lapply(CPDSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE)
))
sixFourSingleCToTMismatchData = lapply(sixFourSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE)
))

```

```{r filter_by_mismatch_pos_and_read_length_1}
CPDPeakCToTMismatchData = lapply(CPDSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS)
})
CPDFirstQPeakCToTMismatchData = lapply(CPDPeakCToTMismatchData, function(x) {
  lapply(x, getFirstPositionQuartile)
})
CPDLastQPeakCToTMismatchData = lapply(CPDPeakCToTMismatchData, function(x) {
  lapply(x, getLastPositionQuartile)
})

sixFourPeakCToTMismatchData = lapply(sixFourSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS)
})
sixFourFirstQPeakCToTMismatchData = lapply(sixFourPeakCToTMismatchData, function(x) {
  lapply(x, getFirstPositionQuartile)
})
sixFourLastQPeakCToTMismatchData = lapply(sixFourPeakCToTMismatchData, function(x) {
  lapply(x, getLastPositionQuartile)
})
```

```{r generate_nucleotide_frequency_data_1}
CPDSingleCToTNucFreq = lapply(CPDSingleCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
CPDFirstQPeakCToTNucFreq = lapply(CPDFirstQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
CPDLastQPeakCToTNucFreq = lapply(CPDLastQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)

sixFourSingleCToTNucFreq = lapply(sixFourSingleCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
sixFourFirstQPeakCToTNucFreq = lapply(sixFourFirstQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
sixFourLastQPeakCToTNucFreq = lapply(sixFourLastQPeakCToTMismatchData, tabulateNucFreqByPosTimepointAndLength)
```

```{r purine_vs_pyrimidine_frequencies_across_read_length, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDSingleCToTNucFreq$rep1, title = "CPD Rep1 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDSingleCToTNucFreq$rep2, title = "CPD Rep2 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourSingleCToTNucFreq$rep1, title = "6-4 Rep1 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourSingleCToTNucFreq$rep2, title = "6-4 Rep2 3' Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```

```{r purine_vs_pyrimidine_frequencies_first_quartile_mismatch_pos, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDFirstQPeakCToTNucFreq$rep1, title = "CPD Rep1 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDFirstQPeakCToTNucFreq$rep2, title = "CPD Rep2 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourFirstQPeakCToTNucFreq$rep1, title = "6-4 Rep1 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourFirstQPeakCToTNucFreq$rep2, title = "6-4 Rep2 1st Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```

```{r purine_vs_pyrimidine_frequencies_last_quartile_mismatch_pos, fig.width = 11, fig.height = 5}
plotNucFreqVsReadLengthBarPlot(CPDLastQPeakCToTNucFreq$rep1, title = "CPD Rep1 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(CPDLastQPeakCToTNucFreq$rep2, title = "CPD Rep2 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))

plotNucFreqVsReadLengthBarPlot(sixFourLastQPeakCToTNucFreq$rep1, title = "6-4 Rep1 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
plotNucFreqVsReadLengthBarPlot(sixFourLastQPeakCToTNucFreq$rep2, title = "6-4 Rep2 4th Quartile Nuc Freq",
                               combineNucleotides = c("AG","CT"), combinedNames = c("Purine","Pyrimidine"))
```

```{r get_file_paths_2}
# Get file paths for the expanded mismatch by read data
expandedCPDMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
expandedCPDMismatchDataFilePaths$rep1 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_CPD_1h_Rep1_mismatches_by_read_3bp_expanded.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_CPD_8h_Rep1_mismatches_by_read_3bp_expanded.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_CPD_24h_Rep1_mismatches_by_read_3bp_expanded.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_CPD_48h_Rep1_mismatches_by_read_3bp_expanded.bed")
)
expandedCPDMismatchDataFilePaths$rep2 = list(
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_CPD_1h_Rep2_mismatches_by_read_3bp_expanded.bed"),
  hr8 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_CPD_8h_Rep2_mismatches_by_read_3bp_expanded.bed"),
  hr24 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_CPD_24h_Rep2_mismatches_by_read_3bp_expanded.bed"),
  hr48 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_CPD_48h_Rep2_mismatches_by_read_3bp_expanded.bed")
)

expandedSixFourMismatchDataFilePaths = list(rep1 = list(), rep2 = list())
expandedSixFourMismatchDataFilePaths$rep1 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_6-4-PP_5min_Rep1_mismatches_by_read_3bp_expanded.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read",
                    "NHF1_6-4-PP_20min_Rep1_mismatches_by_read_3bp_expanded.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_6-4-PP_1h_Rep1_mismatches_by_read_3bp_expanded.bed")
)
expandedSixFourMismatchDataFilePaths$rep2 = list(
  min5 = file.path(deaminationDataDirectory, "mismatches_by_read",
                   "NHF1_6-4-PP_5min_Rep2_mismatches_by_read_3bp_expanded.bed"),
  min20 = file.path(deaminationDataDirectory, "mismatches_by_read",
                    "NHF1_6-4-PP_20min_Rep2_mismatches_by_read_3bp_expanded.bed"),
  hr1 = file.path(deaminationDataDirectory, "mismatches_by_read",
                  "NHF1_6-4-PP_1h_Rep2_mismatches_by_read_3bp_expanded.bed")
)
```

```{r acquire_and_format_data_2}
expandedCPDSingleCToTMismatchData = lapply(expandedCPDMismatchDataFilePaths, function(x) lapply(x, fread))
expandedSixFourSingleCToTMismatchData = lapply(expandedSixFourMismatchDataFilePaths, function(x) lapply(x, fread))
expansionOffset = 3

# Create a subset of single mismatch C>T reads.
expandedCPDSingleCToTMismatchData = lapply(expandedCPDSingleCToTMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))
expandedSixFourSingleCToTMismatchData = lapply(expandedSixFourSingleCToTMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))

# Simplify the data
expandedCPDSingleCToTMismatchData = lapply(expandedCPDSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE, expansionOffset = expansionOffset)
))
expandedSixFourSingleCToTMismatchData = lapply(expandedSixFourSingleCToTMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE, expansionOffset = expansionOffset)
))

# Further filter data by read position based on length
expandedCPDSingleCToTMismatchData = lapply(expandedCPDSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS, expansionOffset)
})

expandedSixFourSingleCToTMismatchData = lapply(expandedSixFourSingleCToTMismatchData, function(x) {
  lapply(x, filterMismatchesByPositionAndReadLength, HUMAN_THREE_PRIME_POS_CONSTRAINTS, expansionOffset)
})
```

```{r orient_data_to_mismatch_positions}
threePrimeMismatchAnchoredCPDData = lapply(expandedCPDSingleCToTMismatchData, function(x) {
  lapply(x, orientDataToMismatch, THREE_PRIME, expansionOffset)
})
fivePrimeMismatchAnchoredCPDData = lapply(expandedCPDSingleCToTMismatchData, function(x) {
  lapply(x, orientDataToMismatch, FIVE_PRIME, expansionOffset)
})

threePrimeMismatchAnchoredSixFourData = lapply(expandedSixFourSingleCToTMismatchData, function(x) {
  lapply(x, orientDataToMismatch, THREE_PRIME, expansionOffset)
})
fivePrimeMismatchAnchoredSixFourData = lapply(expandedSixFourSingleCToTMismatchData, function(x) {
  lapply(x, orientDataToMismatch, FIVE_PRIME, expansionOffset)
})
```

```{r generate_nucleotide_frequency_data_2}
threePrimeMismatchCPDNucFreq = lapply(threePrimeMismatchAnchoredCPDData, tabulateNucFreqByPosTimepointAndLength)
fivePrimeMismatchCPDNucFreq = lapply(fivePrimeMismatchAnchoredCPDData,
                                     tabulateNucFreqByPosTimepointAndLength, posType = FIVE_PRIME)

threePrimeMismatchSixFourNucFreq = lapply(threePrimeMismatchAnchoredSixFourData,
                                          tabulateNucFreqByPosTimepointAndLength)
fivePrimeMismatchSixFourNucFreq = lapply(fivePrimeMismatchAnchoredSixFourData,
                                         tabulateNucFreqByPosTimepointAndLength, posType = FIVE_PRIME)
```

```{r CPD_mismatch__anchored_nucleotide_frequencies, fig.width = 11, fig.height = 7}
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchCPDNucFreq$rep1, 
                               title = "CPD 3' Anchored Nucleotide Frequencies Rep1",
                               secondaryYAxisLabel = "5' Cut Site Distance",
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchCPDNucFreq$rep2, 
                               title = "CPD 3' Anchored Nucleotide Frequencies Rep2",
                               secondaryYAxisLabel = "5' Cut Site Distance",
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)

plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchCPDNucFreq$rep1, 
                               title = "CPD 5' Anchored Nucleotide Frequencies Rep1",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchCPDNucFreq$rep2, 
                               title = "CPD 5' Anchored Nucleotide Frequencies Rep2",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
```

```{r CPD_mismatch_anchored_purine_frequencies, fig.width = 11, fig.height = 7}
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchCPDNucFreq$rep1, 
                               title = "CPD 3' Anchored Purine Frequencies Rep1",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "5' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchCPDNucFreq$rep2, 
                               title = "CPD 3' Anchored Purine Frequencies Rep2",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "5' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)

plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchCPDNucFreq$rep1, 
                               title = "CPD 5' Anchored Purine Frequencies Rep1",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchCPDNucFreq$rep2, 
                               title = "CPD 5' Anchored Purine Frequencies Rep2",
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
```