### 2022-05-16 R/NHF1DeaminationAnchoredExcisionSiteDataGeneration.rmd
Notebook OVERHAUL
- Things were way too messy... They could still be better, but this is definitely a step up without spending an exorbitant amount of time restructuring.
- Individual analysis files have been split into two parts (where relevant): Data generation and figure generation. This should allow for more frequent iterating over the specifics of figures.
- File path generation is now standardized using the new UsefulFileSystemFunctions script
- Implemented general standardizations to parameters that are likely to change as we introduce more data. (E.g. cell type and timepoints).

---
title: "NHF1 Deamination Anchored Excision Site Data Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10
cellType = "NHF1"
CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
sixFourTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)

opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5,
               fig.path = paste0(file.path("deamination_analysis_output", cellType,
                                           "deamination_anchored_excision_site_figures"),
                                 .Platform$file.sep))
if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", cellType)
mismatchesByReadDirectory = file.path(deaminationDataDirectory, "mismatches_by_read")
deaminationExcisionSiteDataDirectory = file.path(deaminationDataDirectory, "excision_site_analysis")
dir.create(deaminationExcisionSiteDataDirectory, showWarnings = FALSE, recursive = TRUE)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationParseAndFilter.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r read_expanded_mismatch_data}
# Get file paths for the expanded mismatch by read data
CPDMismatchData = list(rep1 = CPDTimepoints,
                       rep2 = CPDTimepoints)
for (i in 1:length(CPDMismatchData)) {
  for (j in 1:length(CPDMismatchData[[i]])) {
    CPDMismatchData[[i]][[j]] = fread(generateFilePath(
      mismatchesByReadDirectory, dataTypeStrings$mismatchesByRead, ".bed",
      cellType = cellType, lesion = "CPD", timepoint = names(CPDMismatchData[[i]])[j],
      repitition = names(CPDMismatchData)[i], expansionNum = expansionOffset
      ))
  }
}

sixFourMismatchData = list(rep1 = sixFourTimepoints,
                           rep2 = sixFourTimepoints)
for (i in 1:length(sixFourMismatchData)) {
  for (j in 1:length(sixFourMismatchData[[i]])) {
    sixFourMismatchData[[i]][[j]] = fread(generateFilePath(
      mismatchesByReadDirectory, dataTypeStrings$mismatchesByRead, ".bed",
      cellType = cellType, lesion = "6-4-PP", timepoint = names(sixFourMismatchData[[i]])[j],
      repitition = names(sixFourMismatchData)[i], expansionNum = expansionOffset
      ))
  }
}
```

```{r generate_simplified_and_filtered_mismatch_data}
# Filter the reads by length and mismatch type (single C>T).
CPDMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))
sixFourMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) filterResults(y[V5 == "C>T"], minReadLength = 23, maxReadLength = 31)
))

# Simplify the data
CPDMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE, expansionOffset = expansionOffset)
))
sixFourMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, TRUE, expansionOffset = expansionOffset)
))

# Determine the peak positions to filter on for each read length in each data set.
CPDPeakPositions = lapply(CPDMismatchData, function(x) lapply(
  x, getPeakPositionsByReadLength, expansionOffset
))
sixFourPeakPositions = lapply(sixFourMismatchData, function(x) lapply(
  x, getPeakPositionsByReadLength, expansionOffset
))

# Further filter data by read position based on length
CPDMismatchData = mapply(function(CPDMismatchDataRep, CPDPeakPositionsRep) mapply(
  filterMismatchesByPositionAndReadLength, CPDMismatchDataRep, CPDPeakPositionsRep,
  expansionOffset, SIMPLIFY = FALSE
), CPDMismatchData, CPDPeakPositions, SIMPLIFY = FALSE)
sixFourMismatchData = mapply(function(sixFourMismatchDataRep, sixFourPeakPositionsRep) mapply(
  filterMismatchesByPositionAndReadLength, sixFourMismatchDataRep, sixFourPeakPositionsRep,
  expansionOffset, SIMPLIFY = FALSE
), sixFourMismatchData, sixFourPeakPositions, SIMPLIFY = FALSE)
```

```{r orient_data_to_mismatch_positions}
# NOTE: The assignments to null are there so that the garbage collector can free up memory, and
#       I can still switch to new variable names for organizational purposes.
threePrimeMismatchAnchoredCPDData = lapply(CPDMismatchData, function(x) {
  lapply(x, orientDataToMismatch, THREE_PRIME, expansionOffset)
})
fivePrimeMismatchAnchoredCPDData = lapply(CPDMismatchData, function(x) {
  lapply(x, orientDataToMismatch, FIVE_PRIME, expansionOffset)
})
CPDMismatchData = NULL

threePrimeMismatchAnchoredSixFourData = lapply(sixFourMismatchData, function(x) {
  lapply(x, orientDataToMismatch, THREE_PRIME, expansionOffset)
})
fivePrimeMismatchAnchoredSixFourData = lapply(sixFourMismatchData, function(x) {
  lapply(x, orientDataToMismatch, FIVE_PRIME, expansionOffset)
})
sixFourMismatchData = NULL
```

```{r generate_and_write_nucleotide_frequency_data}
threePrimeMismatchCPDNucFreq = lapply(threePrimeMismatchAnchoredCPDData, tabulateNucFreqByPosTimepointAndLength)
for (i in seq_along(threePrimeMismatchCPDNucFreq)) {
  fwrite(threePrimeMismatchCPDNucFreq[[i]], sep = '\t', file = generateFilePath(
    directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "CPD", repitition = names(threePrimeMismatchCPDNucFreq)[i],
    strandPolarity = "three_prime", anchored = TRUE
  ))
}

fivePrimeMismatchCPDNucFreq = lapply(fivePrimeMismatchAnchoredCPDData,
                                     tabulateNucFreqByPosTimepointAndLength, posType = FIVE_PRIME)
for (i in seq_along(fivePrimeMismatchCPDNucFreq)) {
  fwrite(fivePrimeMismatchCPDNucFreq[[i]], sep = '\t', file = generateFilePath(
    directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMismatchCPDNucFreq)[i],
    strandPolarity = "five_prime", anchored = TRUE
  ))
}

threePrimeMismatchSixFourNucFreq = lapply(threePrimeMismatchAnchoredSixFourData,
                                          tabulateNucFreqByPosTimepointAndLength)
for (i in seq_along(threePrimeMismatchSixFourNucFreq)) {
  fwrite(threePrimeMismatchSixFourNucFreq[[i]], sep = '\t', file = generateFilePath(
    directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "6-4", repitition = names(threePrimeMismatchSixFourNucFreq)[i],
    strandPolarity = "three_prime", anchored = TRUE
  ))
}

fivePrimeMismatchSixFourNucFreq = lapply(fivePrimeMismatchAnchoredSixFourData,
                                         tabulateNucFreqByPosTimepointAndLength, posType = FIVE_PRIME)
for (i in seq_along(fivePrimeMismatchSixFourNucFreq)) {
  fwrite(fivePrimeMismatchSixFourNucFreq[[i]], sep = '\t', file = generateFilePath(
    directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "6-4", repitition = names(fivePrimeMismatchSixFourNucFreq)[i],
    strandPolarity = "five_prime", anchored = TRUE
  ))
}
```

```{r generate_and_write_sequence_enrichment}
threePrimeMMAnchoredCPDEnrichment = mapply(function(mismatchDataRep, nucFreqRep) {
  setNames(lapply(seq_along(mismatchDataRep), function(i) {
    getSequenceFreqByReadLengthAndPos(mismatchDataRep[[i]],
                                      nucFreqRep[Timepoint == names(mismatchDataRep)[i]],
                                      c("CG","CA","TG","TA"))
  }), names(mismatchDataRep))
}, threePrimeMismatchAnchoredCPDData, threePrimeMismatchCPDNucFreq, SIMPLIFY = FALSE)
for (i in seq_along(threePrimeMMAnchoredCPDEnrichment)) {
  for (j in seq_along(threePrimeMMAnchoredCPDEnrichment[[i]])) {
    fwrite(threePrimeMMAnchoredCPDEnrichment[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceEnrichment, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(threePrimeMMAnchoredCPDEnrichment)[i],
      timepoint = names(threePrimeMMAnchoredCPDEnrichment[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "YR"
    ))
  }
}

fivePrimeMMAnchoredCPDEnrichment = mapply(function(mismatchDataRep, nucFreqRep) {
  setNames(lapply(seq_along(mismatchDataRep), function(i) {
    getSequenceFreqByReadLengthAndPos(mismatchDataRep[[i]],
                                      nucFreqRep[Timepoint == names(mismatchDataRep)[i]],
                                      c("TGG"), posType = FIVE_PRIME)
  }), names(mismatchDataRep))
}, fivePrimeMismatchAnchoredCPDData, fivePrimeMismatchCPDNucFreq, SIMPLIFY = FALSE)
for (i in seq_along(fivePrimeMMAnchoredCPDEnrichment)) {
  for (j in seq_along(fivePrimeMMAnchoredCPDEnrichment[[i]])) {
    fwrite(fivePrimeMMAnchoredCPDEnrichment[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceEnrichment, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMMAnchoredCPDEnrichment)[i],
      timepoint = names(fivePrimeMMAnchoredCPDEnrichment[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG"
    ))
  }
}

threePrimeMMAnchoredSixFourEnrichment = mapply(function(mismatchDataRep, nucFreqRep) {
  setNames(lapply(seq_along(mismatchDataRep), function(i) {
    getSequenceFreqByReadLengthAndPos(mismatchDataRep[[i]],
                                      nucFreqRep[Timepoint == names(mismatchDataRep)[i]],
                                      c("CG","CA","TG","TA"))
  }), names(mismatchDataRep))
}, threePrimeMismatchAnchoredSixFourData, threePrimeMismatchSixFourNucFreq, SIMPLIFY = FALSE)
for (i in seq_along(threePrimeMMAnchoredSixFourEnrichment)) {
  for (j in seq_along(threePrimeMMAnchoredSixFourEnrichment[[i]])) {
    fwrite(threePrimeMMAnchoredSixFourEnrichment[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceEnrichment, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(threePrimeMMAnchoredSixFourEnrichment)[i],
      timepoint = names(threePrimeMMAnchoredSixFourEnrichment[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "YR"
    ))
  }
}

fivePrimeMMAnchoredSixFourEnrichment = mapply(function(mismatchDataRep, nucFreqRep) {
  setNames(lapply(seq_along(mismatchDataRep), function(i) {
    getSequenceFreqByReadLengthAndPos(mismatchDataRep[[i]],
                                      nucFreqRep[Timepoint == names(mismatchDataRep)[i]],
                                      c("TGG"), posType = FIVE_PRIME)
  }), names(mismatchDataRep))
}, fivePrimeMismatchAnchoredSixFourData, fivePrimeMismatchSixFourNucFreq, SIMPLIFY = FALSE)
for (i in seq_along(fivePrimeMMAnchoredSixFourEnrichment)) {
  for (j in seq_along(fivePrimeMMAnchoredSixFourEnrichment[[i]])) {
    fwrite(fivePrimeMMAnchoredSixFourEnrichment[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceEnrichment, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(fivePrimeMMAnchoredSixFourEnrichment)[i],
      timepoint = names(fivePrimeMMAnchoredSixFourEnrichment[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG"
    ))
  }
}
```

```{r generate_and_write_sequence_frequencies}
threePrimeMMAnchoredCPDSeqFreq = lapply(threePrimeMismatchAnchoredCPDData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, querySequences = c("CG","CA","TG","TA"))
  })
})
for (i in seq_along(threePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(threePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fwrite(threePrimeMMAnchoredCPDSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(threePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(threePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "YR"
    ))
  }
}

fivePrimeMMAnchoredCPDSeqFreq = lapply(fivePrimeMismatchAnchoredCPDData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, querySequences = c("TGG"), posType = FIVE_PRIME)
  })
})
for (i in seq_along(fivePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(fivePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fwrite(fivePrimeMMAnchoredCPDSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(fivePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG"
    ))
  }
}

threePrimeMMAnchoredSixFourSeqFreq = lapply(threePrimeMismatchAnchoredSixFourData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, querySequences = c("CG","CA","TG","TA"))
  })
})
for (i in seq_along(threePrimeMMAnchoredSixFourSeqFreq)) {
  for (j in seq_along(threePrimeMMAnchoredSixFourSeqFreq[[i]])) {
    fwrite(threePrimeMMAnchoredSixFourSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(threePrimeMMAnchoredSixFourSeqFreq)[i],
      timepoint = names(threePrimeMMAnchoredSixFourSeqFreq[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "YR"
    ))
  }
}

fivePrimeMMAnchoredSixFourSeqFreq = lapply(fivePrimeMismatchAnchoredSixFourData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, querySequences = c("TGG"), posType = FIVE_PRIME)
  })
})
for (i in seq_along(fivePrimeMMAnchoredSixFourSeqFreq)) {
  for (j in seq_along(fivePrimeMMAnchoredSixFourSeqFreq[[i]])) {
    fwrite(fivePrimeMMAnchoredSixFourSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(fivePrimeMMAnchoredSixFourSeqFreq)[i],
      timepoint = names(fivePrimeMMAnchoredSixFourSeqFreq[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG"
    ))
  }
}
```