### 2022-05-19 R/NHF1GeneralReadSequenceDataGeneration.rmd
R notebook updates
- General updates to existing R notebooks
- Added two new notebooks for general read analysis (Looking at sequence frequencies associated with mispriming).

---
title: "NHF1 General Read Sequence Data Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10
cellType = "NHF1"
CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
sixFourTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", cellType)
generalReadSequenceDataDirectory = file.path(deaminationDataDirectory, "general_read_sequences")
dir.create(generalReadSequenceDataDirectory, showWarnings = FALSE, recursive = TRUE)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r define_all_that_and_a_bag_of_chips_function}
singleFileFullAnalysis = function(lesion, repitition, timepoint) {
  readData = fread(generateFilePath(
    generalReadSequenceDataDirectory, dataTypeStrings$alignedReads, ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repitition, expansionNum = expansionOffset
  ))
  readData = data.table(Read_Length = nchar(readData$V7)-expansionOffset*2, Read_Sequence = readData$V7)
  readData = readData[Read_Length >= 23 & Read_Length <= 31]

  for (sequence in c("TGG","CGA")) {
    seqFreq = getSequenceFreqByReadLengthAndPos(readData, querySequences = c(sequence))
    fwrite(seqFreq, sep = '\t', file = generateFilePath(
      directory = generalReadSequenceDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = lesion, repitition = repitition,
      timepoint = timepoint, sequence = sequence
    ))
  }
}
```

```{r compute_and_write_CPD_sequence_frequency}
for (repitition in c("rep1", "rep2")) {
  for (timepoint in names(CPDTimepoints)) {
    singleFileFullAnalysis("CPD", repitition, timepoint)
  }
}
```

```{r compute_and_write_6-4_sequence_frequency}
for (repitition in c("rep1", "rep2")) {
  for (timepoint in names(sixFourTimepoints)) {
    singleFileFullAnalysis("6-4", repitition, timepoint)
  }
}
```