### 2022-05-20 R/NHF1DeaminationSequenceAndPositionDataGeneration.rmd

---
title: "NHF1 Deamination Sequence and Position Data Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered]
---

```{r setup_chunk}
library(knitr)

cellType = "NHF1"
CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
sixFourTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)
expansionOffset = 0

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}
if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", cellType)
mismatchesByReadDirectory = file.path(deaminationDataDirectory, "mismatches_by_read")
deaminationSequenceAndPositionDataDirectory = file.path(deaminationDataDirectory,"sequence_and_position_analysis")
dir.create(deaminationSequenceAndPositionDataDirectory, recursive = TRUE, showWarnings = FALSE)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationParseAndFilter.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationSequenceAndPositionAnalysis.R"))
```

```{r read_mismatch_data}
# Get file paths for the mismatch by read data
CPDMismatchData = list(rep1 = CPDTimepoints,
                       rep2 = CPDTimepoints)
for (i in 1:length(CPDMismatchData)) {
  for (j in 1:length(CPDMismatchData[[i]])) {
    CPDMismatchData[[i]][[j]] = fread(generateFilePath(
      mismatchesByReadDirectory, dataTypeStrings$mismatchesByRead, ".bed",
      cellType = cellType, lesion = "CPD", timepoint = names(CPDMismatchData[[i]])[j],
      repitition = names(CPDMismatchData)[i], filtering = filtering
      ))
  }
}

sixFourMismatchData = list(rep1 = sixFourTimepoints,
                           rep2 = sixFourTimepoints)
for (i in 1:length(sixFourMismatchData)) {
  for (j in 1:length(sixFourMismatchData[[i]])) {
    sixFourMismatchData[[i]][[j]] = fread(generateFilePath(
      mismatchesByReadDirectory, dataTypeStrings$mismatchesByRead, ".bed",
      cellType = cellType, lesion = "6-4-PP", timepoint = names(sixFourMismatchData[[i]])[j],
      repitition = names(sixFourMismatchData)[i], filtering = filtering
      ))
  }
}
```

```{r generate_and_write_simplified_and_filtered_mismatch_data}
# Perform general filtering on the data
CPDMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) filterResults(y, TRUE, 3, 23, 31)
))
sixFourMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) filterResults(y, TRUE, 3, 23, 31)
))

# Simplify the data
CPDMismatchData = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, includeTrinucleotideContext = TRUE)
))
sixFourMismatchData = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) simplifyTable(y, includeTrinucleotideContext = TRUE)
))

# Write the simplified mismatch data
for (i in 1:length(CPDMismatchData)) {
  for (j in 1:length(CPDMismatchData[[i]])) {
    fwrite(CPDMismatchData[[i]][[j]], sep = '\t',
           file = generateFilePath(
             deaminationSequenceAndPositionDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
             cellType = cellType, lesion = "CPD", timepoint = names(CPDMismatchData[[i]])[j],
             repitition = names(CPDMismatchData)[i], filtering = filtering
           ))
  }
}

for (i in 1:length(sixFourMismatchData)) {
  for (j in 1:length(sixFourMismatchData[[i]])) {
    fwrite(sixFourMismatchData[[i]][[j]], sep = '\t',
           file = generateFilePath(
             deaminationSequenceAndPositionDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
             cellType = cellType, lesion = "6-4", timepoint = names(sixFourMismatchData[[i]])[j],
             repitition = names(sixFourMismatchData)[i], filtering = filtering
           ))
  }
}
```

```{r generate_and_write_mismatch_frequency_z-scores}
# Don't forget to subset by C>T mismatches!
CPDMMFreqZScores = lapply(CPDMismatchData, function(x) lapply(
  x, function(y) {
    getMismatchFrequencyZScoreByPosAndReadLength(y[Mismatch == "C>T"], expansionOffset)
  }
))
sixFourMMFreqZScores = lapply(sixFourMismatchData, function(x) lapply(
  x, function(y) {
    getMismatchFrequencyZScoreByPosAndReadLength(y[Mismatch == "C>T"], expansionOffset)
  }
))

for (i in seq_along(CPDMMFreqZScores)) {
  for (j in seq_along(CPDMMFreqZScores[[i]])) {
    fwrite(CPDMMFreqZScores[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(CPDMMFreqZScores)[i],
      timepoint = names(CPDMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = c("C>T")
    ))
  }
}

for (i in seq_along(sixFourMMFreqZScores)) {
  for (j in seq_along(sixFourMMFreqZScores[[i]])) {
    fwrite(sixFourMMFreqZScores[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(sixFourMMFreqZScores)[i],
      timepoint = names(sixFourMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = c("C>T")
    ))
  }
}
```

```{r generate_and_write_mismatch_position_statistics_by_read_length_and_timepoint}
CPDGroupedStats = list(rep1 = list(), rep2 = list())
CPDGroupedStats$rep1$three_prime = getGroupedPositionStats(CPDMismatchData$rep1, includedTypes = c("C>T"))
CPDGroupedStats$rep2$three_prime = getGroupedPositionStats(CPDMismatchData$rep2, includedTypes = c("C>T"))
CPDGroupedStats$rep1$five_prime =
  getGroupedPositionStats(CPDMismatchData$rep1, includedTypes = c("C>T"), posType = FIVE_PRIME)
CPDGroupedStats$rep2$five_prime =
  getGroupedPositionStats(CPDMismatchData$rep2, includedTypes = c("C>T"), posType = FIVE_PRIME)

sixFourGroupedStats = list(rep1 = list(), rep2 = list())
sixFourGroupedStats$rep1$three_prime = getGroupedPositionStats(sixFourMismatchData$rep1, includedTypes = c("C>T"))
sixFourGroupedStats$rep2$three_prime = getGroupedPositionStats(sixFourMismatchData$rep2, includedTypes = c("C>T"))
sixFourGroupedStats$rep1$five_prime =
  getGroupedPositionStats(sixFourMismatchData$rep1, includedTypes = c("C>T"), posType = FIVE_PRIME)
sixFourGroupedStats$rep2$five_prime =
  getGroupedPositionStats(sixFourMismatchData$rep2, includedTypes = c("C>T"), posType = FIVE_PRIME)

for (i in seq_along(CPDGroupedStats)) {
  for (j in seq_along(CPDGroupedStats[[i]])) {
    fwrite(CPDGroupedStats[[i]][[j]], sep = '\t',
           file = generateFilePath(
             deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
             cellType = cellType, lesion = "CPD", repitition = names(CPDGroupedStats)[i],
             strandPolarity = names(CPDGroupedStats[[i]])[j], filtering = filtering
           ))
  }
}

for (i in seq_along(sixFourGroupedStats)) {
  for (j in seq_along(sixFourGroupedStats[[i]])) {
    fwrite(sixFourGroupedStats[[i]][[j]], sep = '\t',
           file = generateFilePath(
             deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
             cellType = cellType, lesion = "6-4", repitition = names(sixFourGroupedStats)[i],
             strandPolarity = names(sixFourGroupedStats[[i]])[j], filtering = filtering
           ))
  }
}
```