### 2022-06-14 R/GM12878DeaminationSequenceAndPositionFigureGeneration.rmd

---
title: "GM12878 Deamination Sequence and Position Figure Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered]
---

```{r setup_chunk}
library(knitr)

cellType = "GM12878"
cisplatinTimepoints = list(`90min` = NULL)
repNames = c("rep1", "rep2", "all_reps")
minReadLength = 22; maxReadLength = 30

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}
if (length(filtering) > 0) {
  titleCaseFiltering = tools::toTitleCase(gsub('_', ' ', filtering))
} else titleCaseFiltering = character()

opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5, 
               fig.path = paste0(file.path("deamination_analysis_output",cellType,
                                           paste(c(filtering,"deamination_sequence_and_position_figures"), collapse = '_')),
                                 .Platform$file.sep))

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", cellType)
deaminationSequenceAndPositionDataDirectory = file.path(deaminationDataDirectory,"sequence_and_position_analysis")
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationSequenceAndPositionAnalysis.R"))
```

```{r read_simplified_mismatch_data}
# Get file paths for the mismatch by read data
cisplatinMismatchData = list(rep1 = cisplatinTimepoints, rep2 = cisplatinTimepoints, all_reps = cisplatinTimepoints)
for (i in seq_along(cisplatinMismatchData)) {
  for (j in seq_along(cisplatinMismatchData[[i]])) {
    cisplatinMismatchData[[i]][[j]] = fread(generateFilePath(
      deaminationSequenceAndPositionDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
      cellType = cellType, lesion = "cisplatin", timepoint = names(cisplatinMismatchData[[i]])[j],
      repitition = names(cisplatinMismatchData)[i], filtering = filtering
    ))
  }
}
```

```{r cisplatin_mismatch_type_frequency}
for (i in 1:length(cisplatinMismatchData)) {
  for (j in 1:length(cisplatinMismatchData[[i]])) {
    plotMismatchTypeFrequencies(cisplatinMismatchData[[i]][[j]],
                                title = paste(c("cisplatin", titleCaseFiltering, names(cisplatinMismatchData)[i], 
                                                names(cisplatinMismatchData[[i]])[j], "Mismatch Type Frequency"),
                                              collapse = ' '))
  }
}
```

```{r cisplatin_mismatch_pos_frequency}
for (i in 1:length(cisplatinMismatchData)) {
  for (j in 1:length(cisplatinMismatchData[[i]])) {
    plotMismatchPositionFrequencies(cisplatinMismatchData[[i]][[j]], includedTypes = c("C>T"),
                                    title = paste(c("cisplatin", titleCaseFiltering, names(cisplatinMismatchData)[i], 
                                                    names(cisplatinMismatchData[[i]])[j], "C>T Mismatch Positions"),
                                                  collapse = ' '))
    plotMismatchPositionFrequencies(cisplatinMismatchData[[i]][[j]], omittedTypes = c("C>T"),
                                    title = paste(c("cisplatin", titleCaseFiltering, names(cisplatinMismatchData)[i], 
                                                    names(cisplatinMismatchData[[i]])[j], "!C>T Mismatch Positions"),
                                                  collapse = ' '))
  }
}
```

```{r cisplatin_trinucleotide_context_analysis}
for (i in 1:length(cisplatinMismatchData)) {
  for (j in 1:length(cisplatinMismatchData[[i]])) {
    plotTrinucleotideContext(cisplatinMismatchData[[i]][[j]], includedTypes = c("C>T"),
                             title = paste(c("cisplatin", titleCaseFiltering, names(cisplatinMismatchData)[i],
                                             names(cisplatinMismatchData[[i]])[j], "C>T Mismatch Trinuc Context"),
                                           collapse = ' '))
  }
}
```

```{r read_mismatch_frequency_z-score_tables}
cisplatinMMFreqZScores = list(rep1 = cisplatinTimepoints, rep2 = cisplatinTimepoints, all_reps = cisplatinTimepoints)
for (i in seq_along(cisplatinMMFreqZScores)) {
  for (j in seq_along(cisplatinMMFreqZScores[[i]])) {
    cisplatinMMFreqZScores[[i]][[j]] = fread(generateFilePath(
      directory = deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "cisplatin", repitition = names(cisplatinMMFreqZScores)[i],
      timepoint = names(cisplatinMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = c("C>T")
    ))
  }
}
```

```{r cisplatin_mismatch_position_frequencies_by_read_length, fig.width = 9, fig.height = 5}
for (i in seq_along(cisplatinMismatchData)) {
  plotPositionAcrossTimepointAndReadLength(cisplatinMismatchData[[i]], includedTypes = c("C>T"),
                                           title = paste(c("cisplatin", titleCaseFiltering, "C>T Mismatch Pos Freq", repNames[i]),
                                                         collapse = ' '),
                                           zScoreTables = cisplatinMMFreqZScores[[i]])
  plotPositionAcrossTimepointAndReadLength(cisplatinMismatchData[[i]], includedTypes = c("C>T"), posType = FIVE_PRIME,
                                           title = paste(c("cisplatin", titleCaseFiltering, "C>T Mismatch Pos Freq", repNames[i]),
                                                         collapse = ' '),
                                           zScoreTables = cisplatinMMFreqZScores[[i]])
}
```

```{r read_read_length_counts}
cisplatinReadLengthCounts = list(rep1 = cisplatinTimepoints, rep2 = cisplatinTimepoints, all_reps = cisplatinTimepoints)
for (i in seq_along(cisplatinReadLengthCounts)) {
  for (j in seq_along(cisplatinReadLengthCounts[[i]])) {
    cisplatinReadLengthCounts[[i]][[j]] = fread(generateFilePath(
      directory = deaminationSequenceAndPositionDataDirectory, dataTypeStrings$readLengthcounts, ".tsv",
      cellType = cellType, lesion = "cisplatin", repitition = names(cisplatinReadLengthCounts)[i],
      timepoint = names(cisplatinReadLengthCounts[[i]])[j], filtering = filtering
    ))
  }
}
```

```{r cisplatin_read_length_distributions}
for (i in seq_along(cisplatinReadLengthCounts)) {
  plotReadLengthFrequencies(cisplatinReadLengthCounts[[i]],
                            title = paste(c("cisplatin", titleCaseFiltering, "Read Length Freq", repNames[i]), collapse = ' '))
}
```

```{r read_mismatch_position_statistics}
cisplatinGroupedStats = list(rep1 = list(three_prime = NULL, five_prime = NULL),
                             rep2 = list(three_prime = NULL, five_prime = NULL),
                             all_reps = list(three_prime = NULL, five_prime = NULL))
for (i in seq_along(cisplatinGroupedStats)) {
  for (j in seq_along(cisplatinGroupedStats[[i]])) {
    cisplatinGroupedStats[[i]][[j]] = fread(generateFilePath(
      deaminationSequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
      cellType = cellType, lesion = "cisplatin", repitition = names(cisplatinGroupedStats)[i],
      strandPolarity = names(cisplatinGroupedStats[[i]])[j], filtering = filtering
    ))
  }
}
```

```{r cisplatin_position_differences_across_read_lengths, fig.width = 9, fig.height = 5}
for (i in seq_along(cisplatinGroupedStats)) {
  plotGroupedPositionStats(cisplatinGroupedStats[[i]]$three_prime, cisplatinGroupedStats[[i]]$five_prime, POS_DIFF,
                           title = paste(c("cisplatin", titleCaseFiltering, "C>T Mismatch Pos Diff", repNames[i]), collapse = ' '),
                           xAxisBreaks = c(minReadLength:maxReadLength))
}
```