### 2022-07-30 R/5'CutSiteAnalysisFigureGeneration.rmd

---
title: "5'CutSiteAnalysisDataGeneration"
author: "Ben MOrledge-Hampton"
date: '2022-07-14'
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered, TG_filtered]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}

if (length(filtering) > 0) {
  titleCaseFiltering = tools::toTitleCase(gsub('_', ' ', filtering))
} else titleCaseFiltering = character()

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "NHF1", "general_read_sequences")
NHF1ExcisionSiteDataDirectory = file.path(deaminationDataDirectory, "NHF1", "excision_site_analysis")
GM12878GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "GM12878", "general_read_sequences")
DmGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Dm", "general_read_sequences")
arabidopsisGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Arabidopsis", "general_read_sequences")
yeastGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "yeast", "general_read_sequences")
EColiGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "E_coli_WT", "general_read_sequences")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
arabidopsisReadMinMax = c(24, 30)
yeastReadMinMax = c(22, 25)
EColiReadMinMax = c(11, 13)

opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5, 
               fig.path = paste0(file.path("deamination_analysis_output",
                                           paste(c(filtering,"5'_cut_site_analysis"), collapse = '_')),
                                 .Platform$file.sep))
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r define_all_that_and_a_bag_of_chips_function}
cutSitePlotting = function(lesion, repetition, timepoint, cellType, directory,
                           sequencesName, readMinMax, maxPlottingPos, xAxisOffset = 1,
                           backgroundPositions = -9:-4, checkPositions = -1,
                           displayPValue = TRUE) {
  sequenceFrequencyData = fread(generateFilePath(
    directory, dataTypeStrings$sequenceFrequency, ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, sequence = sequencesName, filtering = filtering
  ))

  plotSequenceFrequencies(sequenceFrequencyData[Position <= maxPlottingPos],
                          title = paste(cellType, lesion, titleCaseFiltering, timepoint),
                          secondaryYAxisLabel = "Read Length", yAxisLabel = paste(sequencesName,"Frequency"),
                          showFivePrimeCutSite = TRUE, querySequences = c("333"), expansionOffset = 10,
                          minReadLength = readMinMax[1], maxReadLength = readMinMax[2],
                          backgroundPositions = backgroundPositions, checkPositions = checkPositions,
                          displayPValue = displayPValue, xAxisLabel = "5' Relative Position",
                          xAxisOffset = xAxisOffset, xAxisBreaks = c(-8, 0, 8))
}
```

```{r sequence_frequency_figures, fig.width = 5, fig.height = 7}
cutSitePlotting("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("6-4", "all_reps", "20min", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("cisplatin", "rep1", "90min", "GM12878", GM12878GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "20min", "yeast", yeastGeneralReadSequenceDataDirectory,
                "HYV", yeastReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "ZT2", "Arabidopsis", arabidopsisGeneralReadSequenceDataDirectory,
                "HYV", arabidopsisReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "5min", "E_coli_WT", EColiGeneralReadSequenceDataDirectory,
                "HYV", EColiReadMinMax, 7, 1, NULL, NULL, FALSE)
```