### 2022-08-26 R/CombinedReadLengthMMAnchoredSeqFreqGeneration.rmd

---
title: "Combined Read Length Sequence Frequency Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered, TG_filtered]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")

NHF1DataDirectory = file.path(deaminationDataDirectory, "NHF1")
NHF1MismatchesByReadDataDirectory = file.path(NHF1DataDirectory, "mismatches_by_read")
NHF1SequenceAndPositionDataDirectory = file.path(NHF1DataDirectory, "sequence_and_position_analysis")
NHF1ExcisionSiteDataDirectory = file.path(NHF1DataDirectory, "excision_site_analysis")

arabidopsisDataDirectory = file.path(deaminationDataDirectory, "Arabidopsis")
arabidopsisMismatchesByReadDataDirectory = file.path(arabidopsisDataDirectory, "mismatches_by_read")
arabidopsisSequenceAndPositionDataDirectory = file.path(arabidopsisDataDirectory, "sequence_and_position_analysis")
arabidopsisExcisionSiteDataDirectory = file.path(arabidopsisDataDirectory, "excision_site_analysis")

yeastDataDirectory = file.path(deaminationDataDirectory, "yeast")
yeastMismatchesByReadDataDirectory = file.path(yeastDataDirectory, "mismatches_by_read")
yeastSequenceAndPositionDataDirectory = file.path(yeastDataDirectory, "sequence_and_position_analysis")
yeastExcisionSiteDataDirectory = file.path(yeastDataDirectory, "excision_site_analysis")

GM12878DataDirectory = file.path(deaminationDataDirectory, "GM12878")
GM12878SequenceAndPositionDataDirectory = file.path(GM12878DataDirectory, "sequence_and_position_analysis")

EColiDataDirectory = file.path(deaminationDataDirectory, "E_coli_WT")
EColiExcisionSiteDataDirectory = file.path(EColiDataDirectory, "excision_site_analysis")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
arabidopsisReadMinMax = c(24, 30)
yeastReadMinMax = c(22, 25)
EColiReadMinMax = c(11, 13)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationParseAndFilter.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r make_the_function}
computeMMAnchoredCombinedReadLengthSeqFreq = function(
    lesion, repetition, timepoint, cellType, mismatchesByReadDirectory, seqAndPosDirectory,
    excisionSiteDirectory, queryPattern, sequencesName, readMinMax, mismatchType = "C>T"
) {

  if (mismatchType == "C>T") {
    fileNamingMismatchType = character()
  } else {
    fileNamingMismatchType = mismatchType
  }

  # Get file paths for the expanded mismatch by read data
  mismatchData = fread(generateFilePath(
    mismatchesByReadDirectory, dataTypeStrings$mismatchesByRead, ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint, includedMismatches = fileNamingMismatchType,
    repitition = repetition, expansionNum = expansionOffset, filtering = filtering
  ))

  # Filter the reads by length and mismatch type (single C>T).
  mismatchData = filterResults(mismatchData[V5 == mismatchType], minReadLength = readMinMax[1], maxReadLength = readMinMax[2])

  # Simplify the data
  mismatchData = simplifyTable(mismatchData, TRUE, expansionOffset = expansionOffset)

  # Read mismatch frequency z-scores.
  MMFreqZScore = fread(generateFilePath(
    directory = seqAndPosDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
    cellType = cellType, lesion = lesion, repitition = repetition,
    timepoint = timepoint, filtering = filtering, includedMismatches = mismatchType
  ))

  # Further filter data by read position based on length
  mismatchData = filterMismatchesByPositionAndReadLength(mismatchData, MMFreqZScore, expansionOffset = expansionOffset)

  # Orient data to mismatch positions.
  threePrimeMMAnchoredData = orientDataToMismatch(mismatchData, THREE_PRIME, expansionOffset)
  fivePrimeMMAnchoredData = orientDataToMismatch(mismatchData, FIVE_PRIME, expansionOffset)
  mismatchData = NULL

  # Now that the data has been processed, combine read length information.
  threePrimeMMAnchoredData[, Read_Length := NULL]
  threePrimeMMAnchoredData[, Read_Length := "combined"]
  fivePrimeMMAnchoredData[, Read_Length := NULL]
  fivePrimeMMAnchoredData[, Read_Length := "combined"]

  # Generate and write nucleotide frequency data.
  threePrimeMMAnchoredNucFreq = tabulateNucFreqByPosTimepointAndLength(setNames(list(threePrimeMMAnchoredData), timepoint))
  fwrite(threePrimeMMAnchoredNucFreq, sep = '\t', file = generateFilePath(
    directory = excisionSiteDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv", includedMismatches = fileNamingMismatchType,
    cellType = cellType, lesion = lesion, repitition = repetition, timepoint = timepoint, strandPolarity = "three_prime",
    anchored = TRUE, filtering = filtering, additionalInformation = "combined_read_lengths"
  ))
  fivePrimeMMAnchoredNucFreq = tabulateNucFreqByPosTimepointAndLength(setNames(list(fivePrimeMMAnchoredData), timepoint))
  fwrite(fivePrimeMMAnchoredNucFreq, sep = '\t', file = generateFilePath(
    directory = excisionSiteDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv", includedMismatches = fileNamingMismatchType,
    cellType = cellType, lesion = lesion, repitition = repetition, timepoint = timepoint, strandPolarity = "five_prime",
    anchored = TRUE, filtering = filtering, additionalInformation = "combined_read_lengths"
  ))

  # Generate and write sequence frequency data.
  threePrimeMMAnchoredSeqFreq = getSequenceFreqByReadLengthAndPos(threePrimeMMAnchoredData, queryPattern = queryPattern)
  fwrite(threePrimeMMAnchoredSeqFreq, sep = '\t', file = generateFilePath(
    directory = excisionSiteDirectory, dataTypeStrings$sequenceFrequency, ".tsv", includedMismatches = fileNamingMismatchType,
    cellType = cellType, lesion = lesion, repitition = repetition, timepoint = timepoint, strandPolarity = "three_prime",
    anchored = TRUE, sequence = sequencesName, filtering = filtering, additionalInformation = "combined_read_lengths"
  ))

  fivePrimeMMAnchoredSeqFreq = getSequenceFreqByReadLengthAndPos(fivePrimeMMAnchoredData, queryPattern = queryPattern)
  fwrite(fivePrimeMMAnchoredSeqFreq, sep = '\t', file = generateFilePath(
    directory = excisionSiteDirectory, dataTypeStrings$sequenceFrequency, ".tsv", includedMismatches = fileNamingMismatchType,
    cellType = cellType, lesion = lesion, repitition = repetition, timepoint = timepoint, strandPolarity = "five_prime",
    anchored = TRUE, sequence = sequencesName, filtering = filtering, additionalInformation = "combined_read_lengths"
  ))
}
```

```{r compute_and_write_sequence_frequencies}
HYV = c("ACA","ACC","ACG","ATA","ATC","ATG",
        "CCA","CCC","CCG","CTA","CTC","CTG",
        "TCA","TCC","TCG","TTA","TTC","TTG")
HYVPattern = "^[ACT][CT][ACG]"

computeMMAnchoredCombinedReadLengthSeqFreq(
  "CPD", "all_reps", "1h", "NHF1", NHF1MismatchesByReadDataDirectory, NHF1SequenceAndPositionDataDirectory,
  NHF1ExcisionSiteDataDirectory, HYVPattern, "HYV", humanReadMinMax
)
computeMMAnchoredCombinedReadLengthSeqFreq(
  "CPD", "all_reps", "1h", "NHF1", NHF1MismatchesByReadDataDirectory, NHF1SequenceAndPositionDataDirectory,
  NHF1ExcisionSiteDataDirectory, HYVPattern, "HYV", humanReadMinMax, mismatchType = "CC>TT"
)
```