### 2023-03-04 R/MismatchComparisons.rmd
Create MismatchComparisons.rmd
This script compares statistics on mismatches and exports them to a tsv file.

---
title: "Mismatch Comparisons"
author: "Ben Morledge-Hampton"
date: "2023-02-18"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered, TG_filtered]
---

```{r setup_chunk}
library(knitr)

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")

NHF1DataDirectory = file.path(deaminationDataDirectory, "NHF1")
NHF1MismatchesByReadDataDirectory = file.path(NHF1DataDirectory, "mismatches_by_read")

NHF1CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
```

```{r parse_mismatches_by_read_metadata}

summarizeMismatchesByReadMetadata = function(dataDirectory, cellType, lesion, timepoint, verbose = FALSE) {

  if (verbose) {
    print(paste0("Working with ", lesion, " lesions in ", cellType, " cells after ", timepoint, " of repair:"))
  }

  rawMetadata = fread(generateFilePath(
    file.path(dataDirectory,"mismatches_by_read"), dataTypeStrings$mismatchesByRead,
    ".metadata", cellType = cellType, lesion = lesion, repitition = "all_reps",
    timepoint = timepoint
  ))

  totalAlignedReads = sum(rawMetadata[V1 == "Total_Aligned_Reads:", V2])
  totalMismatches = sum(rawMetadata[V1 == "Mismatches:", V2])

  singleMismatchesByRead = fread(generateFilePath(
    file.path(dataDirectory,"sequence_and_position_analysis"), dataTypeStrings$simplifiedMismatches,
    ".tsv", cellType = cellType, lesion = lesion, repitition = "all_reps",
    timepoint = timepoint
  ))

  filteredMismatches = singleMismatchesByRead[,.N]
  singleCtoT_Mismatches = singleMismatchesByRead[Mismatch == "C>T" ,.N]

  tandemMismatchesByRead = fread(generateFilePath(
    file.path(dataDirectory,"sequence_and_position_analysis"), dataTypeStrings$simplifiedMismatches,
    ".tsv", cellType = cellType, lesion = lesion, repitition = "all_reps",
    timepoint = timepoint, includedMismatches = "CC>TT"
  ))

  CCtoTT_Mismatches = tandemMismatchesByRead[,.N]

  filteredMismatchesProportion = filteredMismatches/totalMismatches

  singleCtoT_FilteredMismatchProportion = singleCtoT_Mismatches/totalMismatches

  singleCtoT_ReadProportion = singleCtoT_Mismatches/totalAlignedReads
  CCtoTT_ReadProportion = CCtoTT_Mismatches/totalAlignedReads
  singleToTandemRatio = singleCtoT_Mismatches/CCtoTT_Mismatches

  if (verbose) {
    print(paste0("  Total Aligned Reads: ", totalAlignedReads))
    print(paste0("  Total Mismatches: ", totalMismatches))
    print(paste0("  Filtered Mismatches: ", filteredMismatches))
    print(paste0("    Proportion of mismatches retained after filtering: ", filteredMismatchesProportion))
    print(paste0("  Single C>T Mismatches: ", singleCtoT_Mismatches))
    print(paste0("    Proportion of filtered mismatches that are C>T: ", singleCtoT_FilteredMismatchProportion))
    print(paste0("    Proportion of reads with a single C>T mismatch: ", singleCtoT_ReadProportion))
    print(paste0("  Tandem CC>TT Mismatches: ", CCtoTT_Mismatches))
    print(paste0("    Proportion of reads with a tandem CC>TT mismatch: ", CCtoTT_ReadProportion))
    print(paste0("    Ratio of single C>T mismatches to tandem CC>TT mismatches: ", singleToTandemRatio))
  }

  return(data.table(Cell_Type = cellType, Lesion = lesion, Timepoint = timepoint,
                    Total_Aligned_Reads = totalAlignedReads, Total_Mismatches = totalMismatches,
                    Filtered_Mismatches = filteredMismatches, Filtered_Mismatches_Proportion = filteredMismatchesProportion,
                    Single_C_to_T_Mismatches = singleCtoT_Mismatches,
                    Single_C_to_T_Filtered_Mismatch_Proportion = singleCtoT_FilteredMismatchProportion,
                    Single_C_to_T_Read_Proportion = singleCtoT_ReadProportion, CC_to_TT_Mismatches = CCtoTT_Mismatches,
                    CC_to_TT_Read_Proportion = CCtoTT_ReadProportion, Single_To_Tandem_Ratio = singleToTandemRatio))

}
```

```{r Mismatch_Comparisons}

results = rbindlist(lapply(names(NHF1CPDTimepoints), function(timepoint) {
  summarizeMismatchesByReadMetadata(NHF1DataDirectory, "NHF1", "CPD", timepoint, TRUE)
}))

fwrite(results, file.path(NHF1DataDirectory, "mismatches_by_read", "NHF1_CPD_mismatch_metadata_summary.tsv"),
       sep = '\t')

```