### 2023-03-15 R/CustomFigureGeneration.rmd
Create CustomFigureGeneration.rmd
This notebook serves as a way to create variations on the figures from the PaperFigureGeneration notebook. (e.g. compressing a plot for a poster).

---
title: "Custom Figure Generation"
author: "Ben Morledge-Hampton"
date: "6/21/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)

NHF1CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
NHF1SixFourTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)
arabidopsisCPDTimepoints = list(`ZT2` = NULL, `ZT14` = NULL)
yeastCPDTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)
GM12878CisplatinTimepoints = list(`90min` = NULL)
CSBCPDTimepoints = list(`1h` = NULL)
XPCCPDTimepoints = list(`1h` = NULL)
EColiCPDTimepoints = list(`5min` = NULL)

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

opts_chunk$set(dev = "png", dpi = 600, fig.width = 7, fig.height = 5, 
               fig.path = paste0(file.path("deamination_analysis_output","custom_figures"),
                                 .Platform$file.sep))

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")

NHF1DataDirectory = file.path(deaminationDataDirectory, "NHF1")
NHF1MismatchesByReadDataDirectory = file.path(NHF1DataDirectory, "mismatches_by_read")
NHF1SequenceAndPositionDataDirectory = file.path(NHF1DataDirectory, "sequence_and_position_analysis")
NHF1ExcisionSiteDataDirectory = file.path(NHF1DataDirectory, "excision_site_analysis")
NHF1GeneralReadSequenceDataDirectory = file.path(NHF1DataDirectory, "general_read_sequences")

arabidopsisDataDirectory = file.path(deaminationDataDirectory, "Arabidopsis")
arabidopsisSequenceAndPositionDataDirectory = file.path(arabidopsisDataDirectory, "sequence_and_position_analysis")
arabidopsisExcisionSiteDataDirectory = file.path(arabidopsisDataDirectory, "excision_site_analysis")
arabidopsisGeneralReadSequenceDataDirectory = file.path(arabidopsisDataDirectory, "general_read_sequences")

yeastDataDirectory = file.path(deaminationDataDirectory, "yeast")
yeastSequenceAndPositionDataDirectory = file.path(yeastDataDirectory, "sequence_and_position_analysis")
yeastExcisionSiteDataDirectory = file.path(yeastDataDirectory, "excision_site_analysis")
yeastGeneralReadSequenceDataDirectory = file.path(yeastDataDirectory, "general_read_sequences")

GM12878DataDirectory = file.path(deaminationDataDirectory, "GM12878")
GM12878SequenceAndPositionDataDirectory = file.path(GM12878DataDirectory, "sequence_and_position_analysis")
GM12878GeneralReadSequenceDataDirectory = file.path(GM12878DataDirectory, "general_read_sequences")

CSBDataDirectory = file.path(deaminationDataDirectory, "CS-B")
CSBSequenceAndPositionDataDirectory = file.path(CSBDataDirectory, "sequence_and_position_analysis")

XPCDataDirectory = file.path(deaminationDataDirectory, "XP-C")
XPCSequenceAndPositionDataDirectory = file.path(XPCDataDirectory, "sequence_and_position_analysis")

EColiDataDirectory = file.path(deaminationDataDirectory, "E_coli_WT")
EColiExcisionSiteDataDirectory = file.path(EColiDataDirectory, "excision_site_analysis")
EColiGeneralReadSequenceDataDirectory = file.path(EColiDataDirectory, "general_read_sequences")
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationParseAndFilter.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationSequenceAndPositionAnalysis.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r read_custom_figure_3_data}
NHF1CPDGroupedStatsAllRepsFiltered = list(three_prime = NULL, five_prime = NULL)
for (i in seq_along(NHF1CPDGroupedStatsAllRepsFiltered)) {
  NHF1CPDGroupedStatsAllRepsFiltered[[i]] = fread(generateFilePath(
    NHF1SequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
    cellType = "NHF1", lesion = "CPD", repitition = "all_reps", filtering = "TGG_filtered",
    strandPolarity = names(NHF1CPDGroupedStatsAllRepsFiltered)[i]
  ))
}

arabidopsisCPDGroupedStatsAllRepsFiltered = list(three_prime = NULL, five_prime = NULL)
for (i in seq_along(arabidopsisCPDGroupedStatsAllRepsFiltered)) {
  arabidopsisCPDGroupedStatsAllRepsFiltered[[i]] = fread(generateFilePath(
    arabidopsisSequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
    cellType = "Arabidopsis", lesion = "CPD", repitition = "all_reps", filtering = "TGG_filtered",
    strandPolarity = names(arabidopsisCPDGroupedStatsAllRepsFiltered)[i]
  ))
}

yeastCPDGroupedStatsAllRepsFiltered = list(three_prime = NULL, five_prime = NULL)
for (i in seq_along(yeastCPDGroupedStatsAllRepsFiltered)) {
  yeastCPDGroupedStatsAllRepsFiltered[[i]] = fread(generateFilePath(
    yeastSequenceAndPositionDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
    cellType = "yeast", lesion = "CPD", repitition = "all_reps", filtering = "TGG_filtered",
    strandPolarity = names(yeastCPDGroupedStatsAllRepsFiltered)[i]
  ))
}
```

```{r figure_3C, fig.width = 8, fig.height = 3}
### Just the 1h timepoint ###
plotGroupedPositionStats(NHF1CPDGroupedStatsAllRepsFiltered$three_prime[Timepoint == "1h"],
                         NHF1CPDGroupedStatsAllRepsFiltered$five_prime[Timepoint == "1h"],
                         POS_DIFF, title = "NHF1 CPD 1h All Reps TGG-Filtered C>T Mismatch Pos Diff",
                         xAxisBreaks = c(22:30), yAxisTitle = "Pos Diff")

plotGroupedPositionStats(arabidopsisCPDGroupedStatsAllRepsFiltered$three_prime[Timepoint == "ZT2"],
                         arabidopsisCPDGroupedStatsAllRepsFiltered$five_prime[Timepoint == "ZT2"],
                         POS_DIFF, title = "Arabidopsis CPD 1h All Reps TGG-Filtered C>T Mismatch Pos Diff",
                           xAxisBreaks = c(24:30), yAxisTitle = "Pos Diff")

plotGroupedPositionStats(yeastCPDGroupedStatsAllRepsFiltered$three_prime[Timepoint == "20min"],
                         yeastCPDGroupedStatsAllRepsFiltered$five_prime[Timepoint == "20min"],
                         POS_DIFF, title = "Yeast CPD 1h All Reps TGG-Filtered C>T Mismatch Pos Diff",
                         xAxisBreaks = c(22:25), yAxisTitle = "Pos Diff")

### All timepoints together ###
# plotGroupedPositionStats(NHF1CPDGroupedStatsAllRepsFiltered$three_prime,
#                          NHF1CPDGroupedStatsAllRepsFiltered$five_prime,
#                          POS_DIFF, title = "NHF1 CPD All Reps TGG-Filtered C>T Mismatch Pos Diff",
#                          xAxisBreaks = c(22:30))
# 
# plotGroupedPositionStats(arabidopsisCPDGroupedStatsAllRepsFiltered$three_prime,
#                          arabidopsisCPDGroupedStatsAllRepsFiltered$five_prime,
#                          POS_DIFF, title = "Arabidopsis CPD All Reps TGG-Filtered C>T Mismatch Pos Diff",
#                          xAxisBreaks = c(24:30))
# 
# plotGroupedPositionStats(yeastCPDGroupedStatsAllRepsFiltered$three_prime,
#                          yeastCPDGroupedStatsAllRepsFiltered$five_prime,
#                          POS_DIFF, title = "Yeast CPD All Reps TGG-Filtered C>T Mismatch Pos Diff",
#                          xAxisBreaks = c(22:25))
```