### 2023-07-19 R/NHF1_TFBS_DataGeneration.Rmd
Created systems for XR-seq and TFBS analysis
- Created a function to format and filter mismatched reads to be more easily associated with transcription factor binding sites (TFBSs). (May isolate this to its own script in the near future)
- Created a function to associate TFBSs with XR-seq read mismatches. (Will probably generalize this soon to work with other genomic features, like nucleosomes.)
- Added R functions to further format and plot the above data.

---
title: "NHF1_TFBS_DataGeneration"
author: "Ben Morledge-Hampton"
date: "2023-07-18"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered, TG_filtered]
---

```{r setup_chunk}
library(knitr)

cellType = "NHF1"
CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
mismatchTypes = list(`C>T` = NULL, `CC>TT` = NULL)
TFBSs = list(`CTCF_known1_DHS` = NULL, `CTCF_known1_non-DHS` = NULL)

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
  filtering = readline("Filtering identifier (Input nothing for no filtering):")
  if (filtering == '') filtering = character()
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
  filtering = params$filtering
  if (filtering == "NONE") filtering = character()
}
if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", cellType)
mismatchesByReadDirectory = file.path(deaminationDataDirectory, "mismatches_by_read")
tfbsDataDirectory = file.path(deaminationDataDirectory,"TFBS_analysis")
dir.create(tfbsDataDirectory, showWarnings = FALSE, recursive = TRUE)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationTFBSAnalysis.R"))
```

```{r generate_and_write_TFBS_relative_counts_data}
cpdTfbsRelativeData = copy(CPDTimepoints)
for (i in 1:length(cpdTfbsRelativeData)) {
  cpdTfbsRelativeData[[i]] = copy(mismatchTypes)
  for (j in 1:length(cpdTfbsRelativeData[[i]])) {
    cpdTfbsRelativeData[[i]][[j]] = copy(TFBSs)
    for (k in 1:length(cpdTfbsRelativeData[[i]][[j]])) {
      cpdTfbsRelativeData[[i]][[j]][[k]] = parseMismatchesByTFBS(fread(
        generateFilePath(mismatchesByReadDirectory, dataTypeStrings$mismatchesByReadTfbs, ".bed",
                         cellType = cellType, lesion = "CPD", repitition = "all_reps", filtering = filtering,
                         timepoint = names(cpdTfbsRelativeData)[i], includedMismatches = names(cpdTfbsRelativeData[[i]])[j],
                         additionalInformation = paste0(names(cpdTfbsRelativeData[[i]][[j]])[k],"_related"))
      ))

      fwrite(cpdTfbsRelativeData[[i]][[j]][[k]], sep = '\t', file = generateFilePath(
        tfbsDataDirectory, dataTypeStrings$tfbsRelativeData, ".tsv",
        cellType = cellType, lesion = "CPD", repitition = "all_reps", filtering = filtering,
        timepoint = names(cpdTfbsRelativeData)[i], includedMismatches = names(cpdTfbsRelativeData[[i]])[j],
        additionalInformation = paste0(names(cpdTfbsRelativeData[[i]][[j]])[k],"_related")
      ))

      fwrite(getTfbsRelativeCounts(cpdTfbsRelativeData[[i]][[j]][[k]], feature = TFBS_RELATIVE_POSITION),
             sep = '\t', file = generateFilePath(
        tfbsDataDirectory, dataTypeStrings$tfbsRelativePositionCounts, ".tsv",
        cellType = cellType, lesion = "CPD", repitition = "all_reps", filtering = filtering,
        timepoint = names(cpdTfbsRelativeData)[i], includedMismatches = names(cpdTfbsRelativeData[[i]])[j],
        additionalInformation = paste0(names(cpdTfbsRelativeData[[i]][[j]])[k],"_related")
      ))

      fwrite(getTfbsRelativeCounts(cpdTfbsRelativeData[[i]][[j]][[k]], feature = TFBS_RELATIVE_THREE_PRIME_CUT_SITE),
             sep = '\t', file = generateFilePath(
        tfbsDataDirectory, dataTypeStrings$tfbsRelativeCutSiteCounts, ".tsv", strandPolarity = THREE_PRIME,
        cellType = cellType, lesion = "CPD", repitition = "all_reps", filtering = filtering,
        timepoint = names(cpdTfbsRelativeData)[i], includedMismatches = names(cpdTfbsRelativeData[[i]])[j],
        additionalInformation = paste0(names(cpdTfbsRelativeData[[i]][[j]])[k],"_related")
      ))

      fwrite(getTfbsRelativeCounts(cpdTfbsRelativeData[[i]][[j]][[k]], feature = TFBS_RELATIVE_FIVE_PRIME_CUT_SITE),
             sep = '\t', file = generateFilePath(
        tfbsDataDirectory, dataTypeStrings$tfbsRelativeCutSiteCounts, ".tsv", strandPolarity = FIVE_PRIME,
        cellType = cellType, lesion = "CPD", repitition = "all_reps", filtering = filtering,
        timepoint = names(cpdTfbsRelativeData)[i], includedMismatches = names(cpdTfbsRelativeData[[i]])[j],
        additionalInformation = paste0(names(cpdTfbsRelativeData[[i]][[j]])[k],"_related")
      ))
    }
  }
}

```