### 2023-07-28 R/CutSiteCorrelationAnalysis.rmd
Create CutSiteCorrelationAnalysis.rmd
An R notebook for examining correlations between cut site distances (3' vs 5', relative to the lesion)

---
title: "5'CutSiteAnalysisDataGeneration"
author: "Ben MOrledge-Hampton"
date: '2022-07-14'
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10
filtering = "TGG_filtered"

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1MismatchesByReadDirectory = file.path(deaminationDataDirectory, "NHF1", "mismatches_by_read")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
arabidopsisReadMinMax = c(24, 30)
yeastReadMinMax = c(22, 25)
EColiReadMinMax = c(11, 13)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationAnchoredExcisionSiteAnalysis.R"))
```

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysis = function(lesion, repetition, timepoint, cellType, directory,
                                      includedMismatches = c("CC>TT")) {
  mismatchesByRead = fread(generateFilePath(
    directory, dataTypeStrings$mismatchesByReadRelationFormatted , ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))
  colnames(mismatchesByRead) = c("Chromosome", "Mismatch_Pos_0", "Mismatch_Pos_1", "Three_Prime_Relative_Mismatch_Pos",
                                 "Mismatch_Type", "Mismatch_Strand", "Read_Length")

  mismatchesByRead[,Three_Prime_Cut_Site_Distance := abs(Three_Prime_Relative_Mismatch_Pos)]
  mismatchesByRead[,Five_Prime_Cut_Site_Distance := Read_Length-abs(Three_Prime_Relative_Mismatch_Pos)+1]

  return(cor.test(mismatchesByRead$Three_Prime_Cut_Site_Distance,
                 mismatchesByRead$Five_Prime_Cut_Site_Distance, method = "pearson"))

  #return(mismatchesByRead[,.(Three_Prime_Cut_Site_Distance, Five_Prime_Cut_Site_Distance)])
}
```

```{r compute_and_write_sequence_frequencies}
print(cutSiteCorrelationAnalysis("CPD", "all_reps", "1h", "NHF1", NHF1MismatchesByReadDirectory))
```