### 2023-07-28 R/CutSiteCorrelationDataGeneration.rmd
Split correlation analysis for data and figure gen
- Utilized new suite of cut site correlation analysis functions
- may need to further stratify this by cell type at some point...

---
title: "5'CutSiteAnalysisDataGeneration"
author: "Ben MOrledge-Hampton"
date: '2022-07-14'
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
---

```{r setup_chunk}
library(knitr)

expansionOffset = 10
filtering = "TGG_filtered"

if (interactive()) {
  bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
} else {
  bioinformaticsDirectory = params$bioinformaticsDirectory
}

if (is.na(bioinformaticsDirectory)) stop("No bioinformatics directory given.")

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1MismatchesByReadDirectory = file.path(deaminationDataDirectory, "NHF1", "mismatches_by_read")
NHF1CutSiteCorrelationDirectory = file.path(deaminationDataDirectory, "NHF1", "cut_site_correlation")
dir.create(NHF1CutSiteCorrelationDirectory, showWarnings = FALSE, recursive = TRUE)
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "CutSiteCorrelationAnalysis.R"))
```

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysisDataGeneration = function(lesion, repetition, timepoint, cellType, inputDirectory, outputDirectory,
                                                    includedMismatches = c("CC>TT")) {
  mismatchesByRead = fread(generateFilePath(
    inputDirectory, dataTypeStrings$mismatchesByReadRelationFormatted , ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))
  fwrite(getPairedCutSiteDistances(mismatchesByRead),  sep = '\t', file = generateFilePath(
    outputDirectory, dataTypeStrings$pairedCutSiteDistances , ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))

}
```

```{r generate_cut_site_correlation_data}
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory)
```