### 2023-09-27 R/FeatureRelativeDataGeneration_VALIDATION.Rmd
Validation of feature relative analysis
- Created test data for validation of the feature relative analysis pipeline, including the FilterAndFormat script.
- Test data is given in the validation/feature_relative_analysis directory and is annotated in validation_explanation.bed to describe expected values.
- Jupyter and R markdown notebooks run the validation data through the pipeline.

---
title: "FeatureRelativeDataGeneration_VALIDATION"
author: "Ben Morledge-Hampton"
date: "2023-09-25"
output: html_document
---

```{r setup_chunk}
bioinformaticsDirectory = choose.dir(caption = "Select Bioinformatics_Projects Directory")
filtering = "TGG_filtered"
validationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data", "validation")
```

```{r source_functions}
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "UsefulFileSystemFunctions.R"))
source(file.path(bioinformaticsDirectory, "deamination_determination", 'R',
                 "DeaminationFeatureRelativeAnalysis.R"))
```

```{r generate_and_write_feature_relative_counts_data}
featureRelativeData = parseMismatchesByFeature(fread(generateFilePath(
  validationDataDirectory, dataTypeStrings$mismatchesByReadRelationFormatted, ".bed",
  cellType = "validation", filtering = filtering, includedMismatches = "C>T",
  additionalInformation = "nucleosome_related"
)))

fwrite(featureRelativeData, sep = '\t', file = generateFilePath(
  validationDataDirectory, dataTypeStrings$featureRelativeData, ".tsv",
  cellType = "validation", filtering = filtering, includedMismatches = "C>T",
  additionalInformation = "nucleosome_related"
))

fwrite(getFeatureRelativeCounts(featureRelativeData, dataType = MEAN_THREE_PRIME_CUT_SITE_DISTANCE,
                                strandAlign = TRUE),
       sep = '\t', na = "NA", file = generateFilePath(
  validationDataDirectory, dataTypeStrings$featureRelativeMeanCutSiteDistance, ".tsv", strandPolarity = THREE_PRIME,
  cellType = "validation", filtering = filtering, includedMismatches = "C>T",
  additionalInformation = "nucleosome_related_strand_aligned"
))

fwrite(getFeatureRelativeCounts(featureRelativeData, dataType = MEAN_FIVE_PRIME_CUT_SITE_DISTANCE,
                                strandAlign = TRUE),
       sep = '\t', na = "NA", file = generateFilePath(
  validationDataDirectory, dataTypeStrings$featureRelativeMeanCutSiteDistance, ".tsv", strandPolarity = FIVE_PRIME,
  cellType = "validation", filtering = filtering, includedMismatches = "C>T",
  additionalInformation = "nucleosome_related_strand_aligned"
))
```